---
title: "objective 1.2"
author: "Joseph Lazarus"
date: "7/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(caret)
library(dplyr)
library(skimr)
```

## Import Data

```{r pressure, echo=FALSE}
df <- read.csv("https://raw.githubusercontent.com/nedeinlein/AppliedStatsProject2/main/data_folder/adult.data.csv", strip.white = TRUE)

# change  " ?" to NA
df <- na_if(df,"?")

#check work
colSums(is.na(df))

#target column is income. Well call it a hit if income is ">50K" and assigning it a 1 else 0 this will help match the logit probabilites. Also no NA's in income so this should work. 
df$income <- ifelse(df$income == ">50K",1,0)
df$income <- as.factor(df$income)

#convert the remaining character columns to factors the lazy way.
df[sapply(df,is.character)] <- lapply(df[sapply(df, is.character)], as.factor)

str(df)


#reorder dataframe 

df <- df %>% select(income, age, fnlwgt, education.num, capital.gain, capital.loss, hours.per.week, workclass, education, marital.status, occupation, relationship, race, sex, native.country)

```

## looking for correlations
```{r}
library(GGally)
ggpairs(df[, 1:7], aes(color = income, alpha = 0.4))


ggcorr(df[, 1:7])
```

```{r}
summary(na.omit(df$income))
```

## Create train test split
```{r}
df <- na.omit(df)
summary(df$income)

test_index <- createDataPartition(df$income, p=.3, list = FALSE)
test <- df[test_index,]

table(test$income)

train <- df[-test_index,]

prop.table(table(train$income))
```

## Dealing with imbalanced data sets methods used over sampling, under sampling, mixture of both

```{r}
library(ROSE)
#perform undersampleing. Essentially reduces the number of observations in the majority class income <50k | 0
#calling '$data' extracts the datafram object nested in the ovun.sample object
train_under <- ovun.sample(income ~ ., data = train, method = "under", seed = 1234)$data

table(train_under$income)

#perform oversampling. Essentially replicates observations from the minority class. Income >50k | 1
train_over <- ovun.sample(income ~ ., data = train, method = "over", seed = 1234)$data

table(train_over$income)

#perform a mix of the two
train_both <- ovun.sample(income ~ ., data = train, method = "both", seed = 1234)$data

table(train_both$income)

#now we have three data sets to compare results to. 
#but which variables should we use?
```
##simple Logistic regression using age and education number and even sampling sampling distribution using mixture of both
```{r}
test_simple <- test 
train_simple <- train_both

model.simple <- glm(income ~  education.num + age, binomial(link = 'logit'), data = train_simple)

#store predictions in object to use later
fit.simple <- predict(model.simple, newdata = test_simple, type = "response")

#store predictions in test_simpledf
test_simple$incomeProbability <- fit.simple

#create a prediction column in test_simple
test_simple["Prediction"] = 0

# if else statement at cutoff 0.5
test_simple$Prediction[test_simple$incomeProbability>0.5] = 1

#turn prediction value into factor
test_simple$Prediction=as.factor(test_simple$Prediction)


#confusion matrix: order  predicted classes, Reference
cm.simple <- confusionMatrix(test_simple$Prediction, test_simple$income)

results.lasso<-prediction(fit.simple, test_over$income)
roc.simple = performance(results.lasso, measure = "tpr", x.measure = "fpr")
```

## Variable selection using undersampled
```{r}
library(glmnet)
dat.train.x <- model.matrix(income~., train_under)
dat.train.y<-as.factor(train_under[,1])
cvfit <- cv.glmnet(dat.train.x, dat.train.y, family = "binomial", type.measure = "class", nlambda = 1000)
plot(cvfit)
coef(cvfit, s = "lambda.min")

coef.u<-coef(cvfit,s='lambda.min',exact=TRUE)
inds<-which(coef.u!=0)
variables.u<-row.names(coef.u)[inds]
variables.u<-variables.u[!(variables.u %in% '(Intercept)')]


#CV misclassification error rate is little below .1
print("CV Error Rate:")
cvfit$cvm[which(cvfit$lambda==cvfit$lambda.min)]

#Optimal penalty
print("Penalty Value:")
cvfit$lambda.min

#For final model predictions go ahead and refit lasso using entire
#data set
finalmodel.u<-glmnet(dat.train.x, dat.train.y, family = "binomial",lambda=cvfit$lambda.min)

```


## Variable selection using oversampled
```{r}
library(glmnet)
dat.train.x <- model.matrix(income~., train_over)
dat.train.y<-train_over[,1]
cvfit <- cv.glmnet(dat.train.x, dat.train.y, family = "binomial", type.measure = "class", nlambda = 1000)
plot(cvfit)
coef(cvfit, s = "lambda.min")
#CV misclassification error rate is little below .1
print("CV Error Rate:")
cvfit$cvm[which(cvfit$lambda==cvfit$lambda.min)]

#Optimal penalty
print("Penalty Value:")
cvfit$lambda.min

#For final model predictions go ahead and refit lasso using entire
#data set
finalmodel.o<-glmnet(dat.train.x, dat.train.y, family = "binomial",lambda=cvfit$lambda.min)

```

## Variable selection using mix of both methods
```{r}
library(glmnet)
dat.train.x <- model.matrix(income~., train_both)
dat.train.y<-train_both[,1]
cvfit <- cv.glmnet(dat.train.x, dat.train.y, family = "binomial", type.measure = "class", nlambda = 1000)
plot(cvfit)
coef(cvfit, s = "lambda.min")
#CV misclassification error rate is little below .1
print("CV Error Rate:")
cvfit$cvm[which(cvfit$lambda==cvfit$lambda.min)]

#Optimal penalty
print("Penalty Value:")
cvfit$lambda.min

#For final model predictions go ahead and refit lasso using entire
#data set
finalmodel.b<-glmnet(dat.train.x, dat.train.y, family = "binomial",lambda=cvfit$lambda.min)
coef
```

#make Predictions using undersampled
```{r}
#all models are being compared to the same test set. This duplicates the test set for the under, over, both, sampling schemes
test_under <- test

dat.test.x <- model.matrix(income ~.,test_under)
#use final model from undersampled scheme
fit.pred.lasso <- predict(finalmodel.u, newx = dat.test.x, type="response")

#store lasso predicictions in test_under df as incomeProbability
test_under$incomeProbability <- fit.pred.lasso

#make prediction column to store the probabilities
test_under["Prediction"] = 0

#if else decesion on income probability columns. cutoff .5
test_under$Prediction[test_under$incomeProbability>0.5] = 1
#convert prediction column to factor
test_under$Prediction=as.factor(test_under$Prediction)


#Confusion matrix: order is -> predicted classes, Reference
cm.under <- confusionMatrix(test_under$Prediction, test_under$income)

results.lasso<-prediction(fit.pred.lasso, test_under$income)
roc.lasso.u = performance(results.lasso, measure = "tpr", x.measure = "fpr")

```

## Make Predictions using oversampled
```{r}
#all models are being compared to the same test set. This duplicates the test set for the under, over, both, sampling schemes
test_over <- test

dat.test.x <- model.matrix(income ~.,test_over)

#use finalmodel.o for oversampled scheme
fit.pred.lasso <- predict(finalmodel.o, newx = dat.test.x, type="response")

#store lasso predicictions in test_over df as incomeProbability
test_over$incomeProbability <- fit.pred.lasso

#make prediction column to store the probabilities
test_over["Prediction"] = 0

#if else decision on income probability columns. cutoff .5
test_over$Prediction[test_over$incomeProbability>0.5] = 1
#convert to factor
test_over$Prediction=as.factor(test_over$Prediction)


#Confusion matrix: order is -> predicted classes, Reference
cm.over <- confusionMatrix(test_over$Prediction, test_over$income)

results.lasso<-prediction(fit.pred.lasso, test_over$income)
roc.lasso.o = performance(results.lasso, measure = "tpr", x.measure = "fpr")

```

## Make Predictions using Mix of both sampling schemes
```{r}
#all models are being compared to the same test set. This duplicates the test set for the under, over, both, sampling schemes
test_both <- test

dat.test.x <- model.matrix(income ~.,test_both)

#use finalmodel.b for mix of both sampling schemes
fit.pred.lasso <- predict(finalmodel.b, newx = dat.test.x, type="response")

#store lasso predicictions in test_both df as incomeProbability
test_both$incomeProbability <- fit.pred.lasso

#make prediction column to store the probabilities
test_both["Prediction"] = 0

#if else decision on income probability columns. cutoff .5
test_both$Prediction[test_both$incomeProbability>0.5] = 1
#convert to factor
test_both$Prediction=as.factor(test_both$Prediction)


#Confusion matrix: order is -> predicted classes, Reference
cm.both <- confusionMatrix(test_both$Prediction, test_both$income)

results.lasso<-prediction(fit.pred.lasso, test_both$income)
roc.lasso.b = performance(results.lasso, measure = "tpr", x.measure = "fpr")

```

##Results of lasso variable selection compared with sampling schemes of Under, Over, Both
```{r}
cm.simple
# accuracy 70.11, sensitive 70.35, Specificity 70.35

cm.under
# accuracy 80.48, sensitive 79.46, Specificity 83.53

cm.over
# accuracy 80.32, sensitive 79.21, Specificity 83.67

cm.both
# accuracy 80.35, sensitive 79.05, Specificity 84.29
```
##ROC Curves
```{r}
library(ROCR)
plot(roc.simple)
plot(roc.lasso.u,col ="purple", add = T)
plot(roc.lasso.o, col ="orange", add = T)
plot(roc.lasso.b, col = "blue", add = T)
legend("bottomright",legend=c("Simple","Under","Over","Mix of both"),col=c("black","purple","orange","dodger blue"),lty=1,lwd=1)

```
